A Standardized Effect Size for Evaluating the Strength of Phylogenetic Signal, and Why Lambda is not Appropriate

### 2020-05-10 Manuscript/Manuscript-6-DCA.tex
6

>>>>

This study is not the first to compare \(\lambda\) and \(\kappa\) for
their ability as statistics to measure phylogenetic signal. Our results
for \(\lambda\) and \(\kappa\) values are consistent with those found in
the simulations performed by Münkemüller et al.~(2012), but that study
investigated type I error rates and statistical power, finding that
\(\lambda\) performed better in both regards, irrespective of species
number in trees. Although not the central focus of their study, the same
tendency for variable \(\lambda\) and consistent \(\kappa\) at
intermediate phylogenetic signal strengths was observed (see Fig. 2,
Münkemüller et al. 2012). Recent work by Molina-Venegas and Rodríguez
(2017) found that \(\kappa\) but not \(lambda\) tended to inflate the
estimate of phylogenetic signal, leading to moderate type I and type II
biases, if polytomic chronograms were used. Their work more thoroughly
addressed previous observations of inflated \(\kappa\) for incompletely
resolved phylogenetic trees (Davies et al. 2012; Münkemüller et al.
2012). An interesting question is whether an inflated \(\kappa\) value
leads to an inflated \(Z_\kappa\) or does a tendency of a particular
tree to inflate estimates of \(\kappa\) also inflate the values in
random permutations of a test, in which case \(Z_\kappa\) is robust to
polytomies? We repeated the analyses in Figure 4, adjusting trees to
have 50\% collapsed nodes, per the technique of Molina-Venegas and
Rodríguez (2017), and found results were consistent (see Supporting
Information). This confirms that any tendency of incompletely resolved
trees to inflate \(\kappa\) as a descriptive statistic does not inflate
\(Z_\kappa\) as an effect size, and because comparison of effect sizes
in a test is a comparison of locations of observed values in their
sampling distributions, if the sampling distributions also shift because
of this tendency, the \(Z_{12}\) test statistic in equation 4 appears to
be robust in spite of unresolved trees. \hfill\break
